Hardware-accelerated Rendering of Intersecting Volumes using Boolean Operations

### 2013-04-21 paper.tex
Moved note about early z-kill in first approach below pseudocode
Broke up introduction to first approach
Introduction for first approach summarizes the approach
Use default font size
Use the 'fancyvrb' package for better verbatim environments
Fully detailed first approach

The first approach uses a Boolean AND operation combined with depth-buffer
masking.  It is mostly GPU-based, only requiring that the Boolean operation be
performed on the CPU.  While there is relatively less work on the CPU, it does
require a significant amount of passes.  The pseudocode for the approach is
shown below.  Note that early Z-kill should prevent fragments that are masked
off from being processed by the fragment shader.

@@

>>>>

The first approach uses a Boolean AND operation combined with depth-buffer
masking.  The basic idea is to mask off everything but the area on the screen
with the intersection, rendering the cubes behind it, the intersection itself,
and then the cubes in front of it.  Then the area of the intersection is masked
off and everything else is drawn depth-sorted.

It is mostly GPU-based, only requiring that the Boolean operation be performed
on the CPU.  While there is relatively less work on the CPU, it does require a
significant amount of passes.

The pseudocode for the approach is shown below.

@@

The first step in this approach is to bind all the textures.  There are two for
the volumes, two for storing the texture coordinates of the back faces of the
cubes, and one for storing results between passes.  The last three textures will
be used with external framebuffers, so they should have a width and height at
least as big as the window.

\begin{Verbatim}[fontsize=\small]
  <texture id="first-volume-texture" file="first.vlb">
  <texture id="second-volume-texture" file="second.vlb">
  <texture id="primary-coords-texture" width="768" height="768">
  <texture id="secondary-coords-texture" width="768" height="768">
  <texture id="results-texture" width="768" height="768">
\end{Verbatim}

Next the volumes behind the intersection are rendered into the results texture.
Because we want the part of the volume behind the intersection we first render
the texture coordinates of the back faces of the cubes into a framebuffer, then
render the back faces of the intersection as per \cite{Kruger03}.  Note in the code
snippet below we only show the first volume, hide where we set the uniform
variables in the shader program for brevity, and give the intersection an ID so
it can be redrawn later.

\begin{Verbatim}[fontsize=\small]
  <cull faces="front" />
  <framebuffer>
    <attachment usage="color" link="primary-coords-texture" />
    <use program="primary-coords-program">
      <instance of="cube-1-group" />
    </use>
  </framebuffer>
  <framebuffer>
    <attachment usage="color" link="results-texture" />
    <use program="first-pass-program">
      <booleanAnd id="intersection" of="cube-1 cube-2" />
    </use>
  </framebuffer>
\end{Verbatim}

From there the intersection itself is rendered.  Because the intersection needs
to alternate between samples from both volumes, we have to render the back face
coordinates into two different textures, and then make them both available to
the shader program that renders the intersection.

\begin{Verbatim}[fontsize=\small]
  <cull faces="front" />
  <framebuffer>
    <attachment usage="color" link="coords-texture-1" />
    <use program="primary-coords-program">
      <instance of="intersection" />
    </use>
  </framebuffer>
  <framebuffer>
    <attachment usage="color" link="coords-texture-2" />
    <use program="secondary-coords-program">
      <instance of="intersection" />
    </use>
  </framebuffer>
  <cull faces="back" />
  <framebuffer>
    <attachment usage="color" link="third-pass-texture" />
    <use program="third-pass-program">
      <uniform name="CoordsTexture1" link="coords-texture-1" />
      <uniform name="CoordsTexture2" link="coords-texture-2" />
      <instance of="intersection" />
    </use>
  </framebuffer>
\end{Verbatim}

After that the parts of the volumes in front of the intersection are rendered to
the results texture.  We have the shader program set the depth of the fragments
generated to the lowest value to effectively mask off that part of the screen.

\begin{Verbatim}[fontsize=\small]
  <cull mode="back" />
  <framebuffer>
    <attachment usage="color" link="coords-texture-1" />
    <use program="primary-coords-program">
      <uniform name="MVPMatrix" usage="modelviewprojection" />
      <instance of="cube-1" />
    </use>
  </framebuffer>
  <framebuffer>
    <attachment usage="color" link="results-texture" />
    <use program="fourth-pass-program">
      <instance of="intersection" />
    </use>
  </framebuffer>
\end{Verbatim}

Finally we render the two volumes again depth-sorted with blending enabled and
the depth function set to \emph{LESS}.  Since on the previous step we masked off
the area of the screen with the intersection, the fragments in those parts will
just be discarded.  In fact, on most modern video cards they will not even be
processed by the fragment shader at all because of early Z-kill.

\begin{Verbatim}[fontsize=\small]
  <blend enabled="true" />
  <depth function="less" />
  <sort>
    <group>
      <framebuffer>
        <attachment usage="color" link="primary-coords-texture" />
        <use program="primary-coords-program">
          <cull mode="front" />
          <instance of="cube-1-group" />
        </use>
      </framebuffer>
      <use program="kruger-program">
        <cull faces="back" />
        <group id="cube-1-group">
          <translate x="-1">
            <cube id="cube-1" />
          </translate>
        </group>
      </use>
    </group>
    <group>
      <framebuffer>
        <attachment usage="color" link="coords-texture-2" />
        <use program="primary-coords-program">
          <cull mode="front" />
          <instance of="cube-2-group" />
        </use>
      </framebuffer>
      <use program="kruger-program">
        <cull mode="back" />
        <group id="cube-2-group">
          <translate x="+1">
            <cube id="cube-2" />
          </translate>
        </group>
      </use>
    </group>
  </sort>
\end{Verbatim}

Note that early Z-kill should prevent fragments that are masked off from being
processed by the fragment shader.
